---
title: "Pinto et al. (NHESS): total water level and photo evidence of coastal flooding"
author: "Tom Corringham and Denali Pinto"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  dev_mode: false
  data_dir: "data"
  figures_dir: "output/figures"
  tables_dir: "output/tables"
  transects_rds: "output/twl_backbone.rds"
  twl_photos_rds: "output/twl_photo_panel.rds"
  headline_predictor: "twl_daytime"
  mop_transects_txt: "CA_v1.1_transect_definitions_SD_County.txt"
  mop_txt_skip: 30
  ib_mop_min: "D0037"
  ib_mop_max: "D0057"
  time_xlim_start: "2018-11-01"
  time_xlim_end: "2021-02-28"
  twl_threshold_for_summary: 3.4
  exclude_mops_min_photos_for_threshold_plots: 2
  ar_catalog_xlsx: "ARcatalog_NCEP_NEW_1948-2023_COMPREHENSIVE_08-07-2023_Ranked_Daily_32.5.xlsx"
  ar_catalog_sheet: 1
  ar_window_max_lag_days: 7
  oni_nc: "oni.nc"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.retina = 2
)

# ---- Reproducibility ----
set.seed(123)

# ---- Reproducibility ----
options(mc.cores = 1, brms.backend = "cmdstanr")
Sys.setenv(
  TZ = "UTC",
  OMP_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1"
)

```

Constants
```{r constants}
exclude_mops_default <- c("D0841")

# Headline predictor for threshold figures, Fieller intervals and Bayesian fits.
hp <- params$headline_predictor
predictor_labels <- c(
  twl = "TWL (m)", runup = "Runup (m)",
  twl_daytime = "TWL-daytime (m)", runup_daytime = "Runup-daytime (m)",
  twl_max = "TWL-max (m)", twl_max_daytime = "TWL-max-daytime (m)")
hp_label <- unname(predictor_labels[hp])
hp_name  <- sub(" \\(m\\)$", "", hp_label)
stopifnot(!is.na(hp_label))

# Optional inputs: sections are skipped when the file is absent.
has_ar  <- file.exists(here::here(params$data_dir, params$ar_catalog_xlsx))
has_oni <- file.exists(here::here(params$data_dir, params$oni_nc))
if (!has_ar)  message("AR catalog not found: atmospheric river section skipped.")
if (!has_oni) message("ONI file not found: ENSO section skipped.")
min_photos_keep_default <- params$exclude_mops_min_photos_for_threshold_plots
```


### Packages
```{r packages}
# ---- Packages ----
suppressPackageStartupMessages({
  library(tidyverse)   # dplyr, ggplot2, tidyr, readr, purrr, stringr, tibble
  library(lubridate)
  library(here)
  library(broom)
  library(gridExtra)
  library(viridis)
  library(DT)
  library(readxl)
  library(lme4)
  library(ggtext)
  library(ncdf4)
  library(posterior)
  library(brms)
  library(grid)        # for unit()/arrow() in f11
})

# ---- Fail fast if ggtext missing (optional since we load it above) ----
if (!requireNamespace("ggtext", quietly = TRUE)) {
  stop(
    "Package 'ggtext' is required but not installed.\n",
    "Install it before knitting:\n",
    "install.packages('ggtext')"
  )
}

# Create output folders
dir.create(here::here(params$figures_dir), recursive = TRUE, showWarnings = FALSE)
dir.create(here::here(params$tables_dir),  recursive = TRUE, showWarnings = FALSE)
```

### Helpers
```{r}

assert_file_exists <- function(path) {
  if (!file.exists(path)) stop("File not found: ", path)
  invisible(TRUE)
}

# -----------------------------
# Helper: region from MOP ID
# -----------------------------
mop_region <- function(mop) {
  case_when(
    mop >= "D0038" & mop <= "D0057" ~ "Imperial Beach",
    mop >= "D0495" & mop <= "D0513" ~ "La Jolla",
    mop %in% c("D0666", "D0667")    ~ "Cardiff State Beach",
    TRUE                           ~ "Other"
  )
}

# -----------------------------
# Photo-count bins (drives point/errorbar color + legend)
# -----------------------------
photo_bin <- function(n) {
  case_when(
    n == 1                    ~ "1",
    n >= 2  & n <= 7          ~ "2–7",
    n >= 8  & n <= 12         ~ "8–12",
    n >= 13 & n <= 20         ~ "13–20",
    n >= 21 & n <= 30         ~ "21–30",
    n >= 31 & n <= 50         ~ "31–50",
    n > 50                    ~ "50+",
    TRUE                      ~ NA_character_
  )
}

region_colors <- c(
  "Imperial Beach"      = "#1f77b4",
  "La Jolla"            = "#ff7f0e",
  "Cardiff State Beach" = "#2ca02c",
  "Other"               = "black"
)

photo_colors <- c(
  "1"     = "grey",
  "2–7"   = "violet",
  "8–12"  = "purple",
  "13–20" = "blue",
  "21–30" = "darkblue",
  "31–50" = "darkred",
  "50+"   = "red"
)

save_gg <- function(plot, filename, width=10, height=6, dpi=300) {
  out <- here::here(params$figures_dir, filename)
  ggsave(out, plot = plot, width = width, height = height, units = "in", dpi = dpi)
  message("Saved: ", out)
  invisible(out)
}

dedupe_mop_date_keep_photo <- function(df) {
  # Ensures one row per mop-date, prioritizing photo==1 if duplicates exist.
  df %>%
    group_by(mop, date) %>%
    arrange(desc(photo)) %>%
    slice(1) %>%
    ungroup()
}

add_twl_max_vars <- function(df) {
  # twl_max: max(twl, twl_1_day_before), keeping non-NA if one is missing
  # twl_max_daytime: max(twl_daytime, twl_1_day_before), keeping non-NA if one is missing
  df %>%
    mutate(
      twl_max = case_when(
        is.na(twl) ~ twl_1_day_before,
        is.na(twl_1_day_before) ~ twl,
        TRUE ~ pmax(twl, twl_1_day_before, na.rm = TRUE)
      ),
      twl_max_daytime = case_when(
        is.na(twl_daytime) ~ twl_1_day_before,
        is.na(twl_1_day_before) ~ twl_daytime,
        TRUE ~ pmax(twl_daytime, twl_1_day_before, na.rm = TRUE)
      )
    )
}

fit_logit_threshold <- function(df, predictor, max_abs_threshold = 50, min_abs_slope = 1e-6) {
  f   <- as.formula(paste0("photo ~ ", predictor))
  mod <- suppressWarnings(glm(f, data = df, family = binomial))

  x <- df[[predictor]]; y <- df$photo
  separated <- min(x[y == 1]) >= max(x[y == 0]) || max(x[y == 1]) <= min(x[y == 0])

  co    <- coef(mod)
  slope <- if (length(co) >= 2) unname(co[2]) else NA_real_
  out   <- list(model = mod, threshold = NA_real_, slope = slope,
                converged = mod$converged, separated = separated,
                n_events = sum(y == 1))

  if (separated)      { out$flag <- "separated";      return(out) }
  if (!mod$converged) { out$flag <- "not_converged";  return(out) }
  if (is.na(slope) || abs(slope) < min_abs_slope) {
    out$flag <- "slope_near_zero"; return(out)
  }
  thr <- unname(-co[1] / slope)
  if (!is.finite(thr) || abs(thr) > max_abs_threshold) {
    out$flag <- "threshold_implausible"; return(out)
  }
  out$threshold <- thr; out$flag <- "ok"; out
}

threshold_se_delta <- function(model) {
  b <- coef(model)
  V <- suppressWarnings(vcov(model))

  # Need intercept + slope
  if (length(b) < 2) return(NA_real_)
  b0 <- b[1]; b1 <- b[2]

  # Guard
  if (any(is.na(c(b0, b1))) || !is.finite(b1) || b1 == 0) return(NA_real_)
  if (any(is.na(V[1:2, 1:2]))) return(NA_real_)

  g <- c(-1 / b1, b0 / (b1^2))
  se <- sqrt(as.numeric(t(g) %*% V[1:2, 1:2] %*% g))

  if (!is.finite(se)) return(NA_real_)
  se
}

mop_to_num <- function(mop) as.numeric(gsub("^D", "", mop))
```

### Read inputs
```{r}
transects_path <- here::here(params$transects_rds)
photos_path    <- here::here(params$twl_photos_rds)
mop_txt_path   <- here::here(params$data_dir, params$mop_transects_txt)

assert_file_exists(transects_path)
assert_file_exists(photos_path)
assert_file_exists(mop_txt_path)

data_transects <- readRDS(transects_path)
df_raw <- readRDS(photos_path)

required_df_raw <- c("mop","date","photo","twl","twl_daytime","runup","runup_daytime",
                     "twl_1_day_before","runup_1_day_before")
missing <- setdiff(required_df_raw, names(df_raw))
if (length(missing) > 0) stop("df_raw missing columns: ", paste(missing, collapse=", "))

stopifnot(all(c("mop","date","twl") %in% names(data_transects)))

mop_ll <- readr::read_delim(
  mop_txt_path,
  delim = "\t",
  col_names = c(
    "Sequence number",
    "Label",
    "Backbeach longitude",
    "Backbeach latitude",
    "Prediction site longitude",
    "Prediction site latitude",
    "Prediction site depth",
    "Shore normal",
    "Shore type flag"
  ),
  skip = params$mop_txt_skip,
  show_col_types = FALSE
)
```
```{r dev-checks, include=FALSE}
dplyr::glimpse(df_raw)
range(data_transects$mop, na.rm = TRUE)
```


## Data preparation
```{r}
df <- df_raw %>%
  filter(!is.na(date), !is.na(mop)) %>%
  mutate(date = as.Date(date)) %>%
  filter(str_detect(mop, "^D\\d+$"))

# Keep only MOPs with >=1 photo anywhere in their time series
df_photos <- df %>%
  group_by(mop) %>%
  filter(any(photo == 1, na.rm = TRUE)) %>%
  ungroup()

# De-duplicate mop-date pairs (if present)
df_photos <- dedupe_mop_date_keep_photo(df_photos)

# Add max predictors
df_photos_max <- df_photos %>%
  add_twl_max_vars()

min_events <- params$exclude_mops_min_photos_for_threshold_plots
preds_all  <- c("twl", "runup", "twl_daytime", "runup_daytime", "twl_max", "twl_max_daytime")

mop_events <- df_photos_max %>%
  tidyr::drop_na(dplyr::all_of(c("photo", preds_all))) %>%
  group_by(mop) %>%
  summarise(n_events = sum(photo == 1), n_days = n(), .groups = "drop")

mops_main <- mop_events %>%
  filter(n_events >= min_events, !mop %in% exclude_mops_default) %>%
  pull(mop)

length(mops_main)

# Optionally drop day-before columns for export artifacts
df_photos_export <- df_photos_max %>%
  select(-twl_1_day_before, -runup_1_day_before)
```

```{r}
# Summary counts used in Table/Results text
n_mops_all   <- dplyr::n_distinct(df$mop)
n_mops_photo <- dplyr::n_distinct(df_photos$mop)
n_days       <- dplyr::n_distinct(df_photos$date)
n_photos     <- sum(df_photos$photo == 1, na.rm = TRUE)

summary_tbl <- tibble(
  n_unique_mops_raw        = n_mops_all,
  n_unique_mops_with_photo = n_mops_photo,
  n_unique_days            = n_days,
  total_photos             = n_photos,
  date_min                 = min(df_photos$date, na.rm = TRUE),
  date_max                 = max(df_photos$date, na.rm = TRUE)
)

print(summary_tbl)
```

# Figure: Photos per MOP (with region categories)
```{r}
# Count photos per MOP and assign location groups
df_mop_counts <- df_photos %>%
  dplyr::group_by(mop) %>%
  dplyr::summarise(total_photos = sum(photo == 1, na.rm = TRUE), .groups = "drop") %>%
  dplyr::mutate(
    mop_num = mop_to_num(mop),
    location = dplyr::case_when(
      mop >= "D0038" & mop <= "D0057" ~ "Imperial Beach",
      mop >= "D0495" & mop <= "D0513" ~ "La Jolla",
      mop %in% c("D0666", "D0667")    ~ "Cardiff State Beach",
      TRUE                            ~ "Other"
    )
  ) %>%
  dplyr::arrange(mop_num) %>%
  dplyr::mutate(mop = factor(mop, levels = mop))

# Consistent manuscript palette (matches your second chunk)
loc_colors <- c(
  "Imperial Beach"       = "#1f77b4",
  "La Jolla"             = "#ff7f0e",
  "Cardiff State Beach"  = "#2ca02c",
  "Other"                = "#4D4D4D"  
)

df_mop_counts <- df_mop_counts %>%
  dplyr::mutate(
    mop_label = paste0(
      "<span style='color:", loc_colors[location], "'>",
      as.character(mop),
      "</span>"
    )
  )

p_photos_per_mop <- ggplot(df_mop_counts, aes(x = mop, y = total_photos, fill = location)) +
  geom_col(width = 0.85) +
  scale_fill_manual(values = loc_colors, drop = FALSE) +
  scale_x_discrete(labels = setNames(gsub("<[^>]+>", "", df_mop_counts$mop_label), df_mop_counts$mop)) +
  labs(
    x = "MOP Transect",
    y = "Total Photos",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.position = c(0.85, 0.95),
    legend.justification = "top",
    legend.margin = margin(r = 20),
    plot.margin = margin(r = 40)
  )

p_photos_per_mop
save_gg(p_photos_per_mop, "f07.png", width = 10, height = 6)
```

# Figure: Monthly photographed incidents
```{r}
df_monthly <- df_photos %>%
  filter(photo == 1) %>%
  mutate(month_num = lubridate::month(date)) %>%
  count(month_num) %>%
  mutate(
    month = factor(month_num,
                   levels = c(10,11,12,1,2,3,4,5,6,7,8,9),
                   labels = c("Oct","Nov","Dec","Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep"))
  )

p_monthly <- ggplot(df_monthly, aes(x = month, y = n, fill = month)) +
  geom_col(show.legend = FALSE) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = NULL, y = "Number of Photographed Incidents") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8)
  )

p_monthly
save_gg(p_monthly, "f04.png", width = 10, height = 6)

mean(df_monthly$n)
var(df_monthly$n)
```
# Figure: Year and month of photographed incidents
```{r}
# ----- Year x Month photographed incidents (Oct–Sep order, viridis palette) -----

data_clean <- df_photos_export  # or df_photos_max / df_photos

# Ensure required columns exist
stopifnot(all(c("date", "photo") %in% names(data_clean)))

# Ensure Date format
data_clean <- data_clean %>%
  mutate(date = as.Date(date))

# Custom month order (October–September)
month_order <- c(
  "October", "November", "December", "January", "February", "March",
  "April", "May", "June", "July", "August", "September"
)

# Prepare counts
yearly_monthly_counts <- data_clean %>%
  mutate(
    year  = lubridate::year(date),
    month = lubridate::month(date, label = TRUE, abbr = FALSE),
    month = factor(as.character(month), levels = month_order)
  ) %>%
  filter(photo == 1) %>%
  count(year, month, name = "n")

p_year_month <- ggplot(yearly_monthly_counts,
                       aes(x = factor(year), y = n, fill = month)) +
  geom_col(position = position_dodge2(preserve = "single")) +
  scale_fill_viridis(discrete = TRUE) +   
  labs(
    x = NULL,
    y = "Number of Photographed Incidents",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = -45, hjust = 0.5, vjust = 0.5, size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(margin = margin(r = 10), size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
  )

p_year_month
save_gg(p_year_month,
        "f05.png",
        width = 10, height = 6)
```


# TWL Time Series with Logistic sigmoid curve
```{r, eval=FALSE}
# Superseded by the Options A/B chunk below, which also draws the balanced documentation threshold.
# Figure: One MOP time series + right panel logistic (D0044) using TWO predictors
mop_id <- "D0044"

# Predictors used in the manuscript and their axis labels / filenames
predictors_fig08_09 <- tribble(
  ~predictor_var,      ~predictor_label,        ~file_stub,
  hp,                   hp_label,                "f09",
  "twl",                "TWL (m)",              "f08"
)

# Filter to IB + the selected MOP once
df_ib <- df_photos_max %>%
  filter(mop >= params$ib_mop_min, mop <= params$ib_mop_max) %>%
  mutate(date = as.Date(date))

df_mop_all <- df_ib %>% filter(mop == mop_id)
stopifnot(nrow(df_mop_all) > 0)

# ---- Shared y-axis limits across BOTH predictors (same scale) ----
y_max_pad <- 0.5
y_min <- suppressWarnings(min(c(df_mop_all[[hp]], df_mop_all$twl), na.rm = TRUE))
y_max <- suppressWarnings(max(c(df_mop_all[[hp]], df_mop_all$twl), na.rm = TRUE))

if (!is.finite(y_min) || !is.finite(y_max)) {
  stop("Could not compute shared y-limits for ", mop_id, ". Check that twl and the headline predictor exist and are numeric.")
}
y_limits <- c(y_min, y_max + y_max_pad)

# ---- Helper to build the two-panel figure for one predictor ----
make_timeseries_logistic_figure <- function(df_mop_all, predictor_var, predictor_label, file_stub) {

  df_mop <- df_mop_all %>%
    filter(!is.na(.data[[predictor_var]]))
  stopifnot(nrow(df_mop) > 0)

  # Fit model: photo ~ predictor
  fit <- fit_logit_threshold(df_mop, predictor_var)
  mod <- fit$model
  thr <- fit$threshold

  # Build logistic curve
  twl_seq <- seq(
    min(df_mop[[predictor_var]], na.rm = TRUE),
    max(df_mop[[predictor_var]], na.rm = TRUE) + 1,
    length.out = 200
  )
  newdata <- tibble(!!predictor_var := twl_seq)

  logistic_curve <- tibble(
    twl  = twl_seq,
    prob = predict(mod, newdata = newdata, type = "response")
  )

  # ---- Legend levels (exactly what you want, in order) ----
  legend_levels <- c(
    "Evidence of impact",
    predictor_label,
    "0.5 probability threshold",
    "Logistic sigmoid curve"
  )

  # ---- Robust named scale values (always 4, always aligned) ----
  color_values <- setNames(
    c("red", "grey40", "forestgreen", "blue"),
    legend_levels
  )

  linetype_values <- setNames(
    c("blank", "solid", "dashed", "dashed"),
    legend_levels
  )

  shape_values <- setNames(
    c(1, NA, NA, NA),
    legend_levels
  )

  # Override aesthetics MUST match length(legend_levels)
  override_shapes    <- c(1, NA, NA, NA)
  override_linetypes <- c("blank", "solid", "dashed", "dashed")
  override_linewidth <- c(0, 0.8, 0.8, 0.8)

  # Left panel: time series (legend lives HERE)
  p_left <- ggplot(df_mop, aes(x = date)) +

    # TWL line -> legend item
    geom_line(
      aes(
        y = .data[[predictor_var]],
        color = predictor_label,
        linetype = predictor_label
      ),
      linewidth = 0.7
    ) +

    # Impact points (red open circles) -> legend item
    geom_point(
      data = df_mop %>% filter(photo == 1),
      aes(
        y = .data[[predictor_var]],
        color = "Evidence of impact",
        shape = "Evidence of impact",
        linetype = "Evidence of impact"
      ),
      size = 2.6, stroke = 1
    ) +

    # Dummy logistic curve legend key (blue dashed) — invisible but appears in legend
    geom_line(
      data = logistic_curve,
      aes(
        x = df_mop$date[1], y = twl,
        color = "Logistic sigmoid curve",
        linetype = "Logistic sigmoid curve"
      ),
      linewidth = 0.7, alpha = 0
    ) +

    # Dummy threshold legend key (green dashed) — invisible but appears in legend
    geom_hline(
      aes(
        yintercept = thr,
        color = "0.5 probability threshold",
        linetype = "0.5 probability threshold"
      ),
      linewidth = 0.7, alpha = 0
    ) +

    scale_x_date(
      limits = as.Date(c(params$time_xlim_start, params$time_xlim_end)),
      date_labels = "%b %Y",
      date_breaks = "6 months",
      expand = c(0, 0)
    ) +
    scale_y_continuous(limits = y_limits, expand = c(0, 0)) +
    labs(x = NULL, y = predictor_label, color = NULL, linetype = NULL, shape = NULL) +

    # Unified legend scales (robust)
    scale_color_manual(values = color_values, breaks = legend_levels, drop = FALSE) +
    scale_linetype_manual(values = linetype_values, breaks = legend_levels, drop = FALSE) +
    scale_shape_manual(values = shape_values, breaks = legend_levels, drop = FALSE) +

    guides(
      color = guide_legend(
        ncol = 1,
        override.aes = list(
          shape     = override_shapes,
          linetype  = override_linetypes,
          linewidth = override_linewidth,
          alpha     = rep(1, length(legend_levels))
        )
      ),
      linetype = "none",
      shape    = "none"
    ) +

    theme_minimal() +
    theme(
      axis.text = element_text(size = 11, color = "black"),
      axis.title.y = element_text(size = 12, margin = margin(r = 10)),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
      panel.grid = element_blank(),

      # legend inside, top-center of time series
      legend.position = c(0.5, 0.95),
      legend.justification = c(0.5, 1),
      legend.background = element_rect(fill = scales::alpha("white", 0.75), color = NA),
      legend.key = element_rect(fill = NA, color = NA),
      legend.text = element_text(size = 10)
    )

  # Right panel: 0/1 impacts + logistic curve in probability space (no legend)
  p_right <- ggplot() +
    geom_point(
      data = df_mop %>% filter(photo == 0),
      aes(x = 0, y = .data[[predictor_var]]),
      shape = 1, size = 2.6, stroke = 1, color = "grey60"
    ) +
    geom_point(
      data = df_mop %>% filter(photo == 1),
      aes(x = 1, y = .data[[predictor_var]]),
      shape = 1, size = 2.6, stroke = 1, color = "red"
    ) +
    geom_line(
      data = logistic_curve,
      aes(x = prob, y = twl),
      linetype = "dashed", linewidth = 0.7, color = "blue"
    ) +
    geom_hline(
      yintercept = thr,
      linetype = "dashed", linewidth = 0.7, color = "forestgreen"
    ) +
    scale_x_continuous(
      breaks = c(0, 1),
      labels = c("No evidence\nof impact", "Evidence\nof impact"),
      limits = c(-0.5, 1.5)
    ) +
    scale_y_continuous(limits = y_limits, expand = c(0, 0)) +
    labs(x = NULL, y = predictor_label) +
    theme_minimal() +
    theme(
      axis.text = element_text(size = 11, color = "black"),
      axis.title.y = element_text(size = 12, margin = margin(r = 10)),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
      panel.grid = element_blank(),
      legend.position = "none"
    )

  g <- grid.arrange(p_left, p_right, ncol = 2, widths = c(2.5, 0.9))

  # Save
  out_path <- here::here(params$figures_dir, paste0(file_stub, ".png"))
  ggsave(out_path, plot = g, width = 12, height = 6, units = "in", dpi = 300)
  if (isTRUE(params$dev_mode)) message("Saved: ", out_path)

  invisible(g)
}

# ---- Build and save BOTH manuscript figures ----
plots <- predictors_fig08_09 %>%
  mutate(
    plot = pmap(
      list(predictor_var, predictor_label, file_stub),
      ~make_timeseries_logistic_figure(
        df_mop_all = df_mop_all,
        predictor_var = ..1,
        predictor_label = ..2,
        file_stub = ..3
      )
    )
  )

# Print them into the knitted output (order matches tribble rows)
plots$plot[[1]]
plots$plot[[2]]
```

edit Sept 26, 2026
# TWL Time Series with Logistic sigmoid curve (Options A and B)
```{r}
# Option A ("full"):   left panel shows the full record; right panel shows all days.
# Option B ("window"): left panel zoomed to params$time_xlim_*; right panel points
#                      restricted to the same window.
# In BOTH options the logistic fit and 0.5 threshold use the FULL record.
mop_id <- "D0044"

predictors_fig08_09 <- tribble(
  ~predictor_var,      ~predictor_label,        ~file_stub,
  hp,                   hp_label,                "f09",
  "twl",                "TWL (m)",              "f08"
)

view_options <- tribble(
  ~view,     ~suffix,       ~fig_width,
  "full",    "_A_full",     14,
  "window",  "_B_window",   12
)

df_ib <- df_photos_max %>%
  filter(mop >= params$ib_mop_min, mop <= params$ib_mop_max) %>%
  mutate(date = as.Date(date))

df_mop_all <- df_ib %>% filter(mop == mop_id)
stopifnot(nrow(df_mop_all) > 0)

# Shared y-limits across both predictors AND both views (full record)
y_max_pad <- 0.5
y_min <- suppressWarnings(min(c(df_mop_all[[hp]], df_mop_all$twl), na.rm = TRUE))
y_max <- suppressWarnings(max(c(df_mop_all[[hp]], df_mop_all$twl), na.rm = TRUE))
if (!is.finite(y_min) || !is.finite(y_max)) {
  stop("Could not compute shared y-limits for ", mop_id, ".")
}
y_limits <- c(y_min, y_max + y_max_pad)

win <- as.Date(c(params$time_xlim_start, params$time_xlim_end))

make_timeseries_logistic_figure <- function(df_mop_all, predictor_var, predictor_label,
                                            file_stub, view = c("full", "window"),
                                            suffix = "", fig_width = 12) {
  view <- match.arg(view)

  df_mop <- df_mop_all %>% filter(!is.na(.data[[predictor_var]]))
  stopifnot(nrow(df_mop) > 0)

  # ---- Fit on the FULL record (unchanged across views) ----
  fit <- fit_logit_threshold(df_mop, predictor_var)
  mod <- fit$model
  thr <- fit$threshold
  # Balanced documentation threshold: fitted probability equals the overall photo frequency
  balanced_thr <- thr + qlogis(mean(df_mop$photo)) / fit$slope

  twl_seq <- seq(
    min(df_mop[[predictor_var]], na.rm = TRUE),
    max(df_mop[[predictor_var]], na.rm = TRUE) + 1,
    length.out = 200
  )
  newdata <- tibble(!!predictor_var := twl_seq)
  logistic_curve <- tibble(
    twl  = twl_seq,
    prob = predict(mod, newdata = newdata, type = "response")
  )

  # ---- Data actually DISPLAYED (differs by view) ----
  # Option A drops 2010 from the DISPLAY only; the fit above still uses it.
  if (view == "window") {
    df_plot     <- df_mop %>% filter(date >= win[1], date <= win[2])
    x_lims      <- win
    date_breaks <- "6 months"
    date_labels <- "%b %Y"
  } else {
    df_plot     <- df_mop %>% filter(lubridate::year(date) != 2010)
    x_lims      <- range(df_plot$date, na.rm = TRUE)
    date_breaks <- "1 year"
    date_labels <- "%Y"
  }
  stopifnot(nrow(df_plot) > 0)

  message(sprintf("%s | %s | impacts displayed: %d of %d in full record",
                  predictor_var, view,
                  sum(df_plot$photo == 1), sum(df_mop$photo == 1)))

  # ---- Legend setup ----
  legend_levels <- c("Evidence of impact", predictor_label,
                     "Documentation threshold", "Balanced documentation threshold",
                     "Logistic sigmoid curve")
  color_values    <- setNames(c("red", "grey40", "forestgreen", "darkorange", "blue"), legend_levels)
  linetype_values <- setNames(c("blank", "solid", "dashed", "longdash", "dashed"), legend_levels)
  shape_values    <- setNames(c(1, NA, NA, NA, NA), legend_levels)

  # ---- Left panel ----
  p_left <- ggplot(df_plot, aes(x = date)) +
    geom_line(
      aes(y = .data[[predictor_var]], color = predictor_label, linetype = predictor_label),
      linewidth = 0.7
    ) +
    geom_point(
      data = df_plot %>% filter(photo == 1),
      aes(y = .data[[predictor_var]], color = "Evidence of impact",
          shape = "Evidence of impact", linetype = "Evidence of impact"),
      size = 2.6, stroke = 1
    ) +
    geom_line(
      data = logistic_curve,
      aes(x = x_lims[1], y = twl, color = "Logistic sigmoid curve",
          linetype = "Logistic sigmoid curve"),
      linewidth = 0.7, alpha = 0
    ) +
    geom_hline(
      aes(yintercept = thr, color = "Documentation threshold",
          linetype = "Documentation threshold"),
      linewidth = 0.7, alpha = 0
    ) +
    geom_hline(
      aes(yintercept = balanced_thr, color = "Balanced documentation threshold", linetype = "Balanced documentation threshold"),
      linewidth = 0.7, alpha = 0
    ) +
    scale_x_date(limits = x_lims, date_labels = date_labels,
                 date_breaks = date_breaks, expand = c(0, 0)) +
    scale_y_continuous(limits = y_limits, expand = c(0, 0)) +
    labs(x = NULL, y = predictor_label, color = NULL, linetype = NULL, shape = NULL) +
    scale_color_manual(values = color_values, breaks = legend_levels, drop = FALSE) +
    scale_linetype_manual(values = linetype_values, breaks = legend_levels, drop = FALSE) +
    scale_shape_manual(values = shape_values, breaks = legend_levels, drop = FALSE) +
    guides(
      color = guide_legend(
        ncol = 1,
        override.aes = list(
          shape     = c(1, NA, NA, NA, NA),
          linetype  = c("blank", "solid", "dashed", "longdash", "dashed"),
          linewidth = c(0, 0.8, 0.8, 0.8, 0.8),
          alpha     = rep(1, length(legend_levels))
        )
      ),
      linetype = "none", shape = "none"
    ) +
    theme_minimal() +
    theme(
      axis.text = element_text(size = 11, color = "black"),
      axis.title.y = element_text(size = 12, margin = margin(r = 10)),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
      panel.grid = element_blank(),
      legend.position = c(0.5, 0.99),
      legend.justification = c(0.5, 1),
      legend.background = element_rect(fill = scales::alpha("white", 0.75), color = NA),
      legend.key = element_rect(fill = NA, color = NA),
      legend.text = element_text(size = 10)
    )

  # ---- Right panel: points from df_plot, curve/threshold from full-record fit ----
  p_right <- ggplot() +
    geom_point(
      data = df_plot %>% filter(photo == 0),
      aes(x = 0, y = .data[[predictor_var]]),
      shape = 1, size = 2.6, stroke = 1, color = "grey60"
    ) +
    geom_point(
      data = df_plot %>% filter(photo == 1),
      aes(x = 1, y = .data[[predictor_var]]),
      shape = 1, size = 2.6, stroke = 1, color = "red"
    ) +
    geom_line(
      data = logistic_curve, aes(x = prob, y = twl),
      linetype = "dashed", linewidth = 0.7, color = "blue"
    ) +
    geom_hline(yintercept = thr, linetype = "dashed", linewidth = 0.7, color = "forestgreen") +
    geom_hline(yintercept = balanced_thr, linetype = "longdash", linewidth = 0.7, color = "darkorange") +
    scale_x_continuous(breaks = c(0, 1), labels = c("No evidence\nof impact", "Evidence\nof impact"),
                       limits = c(-0.6, 1.6)) +
    scale_y_continuous(limits = y_limits, expand = c(0, 0)) +
    labs(x = NULL, y = predictor_label) +
    theme_minimal() +
    theme(
      axis.text = element_text(size = 11, color = "black"),
      axis.text.x = element_text(size = 9, color = "black"),
      axis.title.y = element_text(size = 12, margin = margin(r = 10)),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
      panel.grid = element_blank(),
      legend.position = "none"
    )

  g <- gridExtra::arrangeGrob(p_left, p_right, ncol = 2, widths = c(2.5, 1.2))

  out_path <- here::here(params$figures_dir, paste0(file_stub, suffix, ".png"))
  ggsave(out_path, plot = g, width = fig_width, height = 6, units = "in", dpi = 300)
  if (isTRUE(params$dev_mode)) message("Saved: ", out_path)

  g
}

# ---- Build all four figures (2 predictors x 2 views) ----
plots <- tidyr::crossing(predictors_fig08_09, view_options) %>%
  mutate(
    plot = pmap(
      list(predictor_var, predictor_label, file_stub, view, suffix, fig_width),
      ~make_timeseries_logistic_figure(
        df_mop_all      = df_mop_all,
        predictor_var   = ..1,
        predictor_label = ..2,
        file_stub       = ..3,
        view            = ..4,
        suffix          = ..5,
        fig_width       = ..6
      )
    )
  )

# Print into the knitted output
for (i in seq_len(nrow(plots))) {
  grid::grid.newpage()
  grid::grid.draw(plots$plot[[i]])
}
```



### Logistic threshold for all MOPs and predictors
```{r}
predictors <- c("twl", "runup", "twl_daytime", "runup_daytime", "twl_max", "twl_max_daytime")

results <- map_dfr(sort(unique(df_photos_max$mop)), function(mop_id) {
  df_m <- df_photos_max %>% filter(mop == mop_id)

  total_photos <- sum(df_m$photo == 1, na.rm = TRUE)
  # Need both outcome classes
  if (nrow(df_m) < 10 || n_distinct(df_m$photo) < 2) {
    return(tibble())
  }

  map_dfr(predictors, function(p) {
    # Require predictor to exist and not be all NA
    if (!(p %in% names(df_m)) || all(is.na(df_m[[p]]))) {
      return(tibble())
    }

    df_mm <- df_m %>% filter(!is.na(.data[[p]]), !is.na(photo))

    if (nrow(df_mm) < 10 || n_distinct(df_mm$photo) < 2) {
      return(tibble())
    }

    fit <- tryCatch(fit_logit_threshold(df_mm, p), error = function(e) NULL)
    if (is.null(fit)) return(tibble())

    mod <- fit$model
    thr <- fit$threshold

    # Extract inference
    td <- broom::tidy(mod)
    pval <- td %>% filter(term == p) %>% pull(p.value)
    coef_value <- td %>% filter(term == p) %>% pull(estimate)

    se_thr <- threshold_se_delta(mod)

    if (is.na(thr) || is.na(se_thr)) {
      lb <- NA_real_
      ub <- NA_real_
    } else {
      lb <- thr - 1.96 * se_thr
      ub <- thr + 1.96 * se_thr
    }

    tibble(
      mop = mop_id,
      predictor = p,
      total_photos = total_photos,
      threshold = thr,
      threshold_se = se_thr,
      lower_95 = lb,
      upper_95 = ub,
      coefficient = coef_value,
      slope = fit$slope,
      fit_flag = fit$flag,
      n_events  = fit$n_events,
      converged = fit$converged,
      separated = fit$separated,
      p_value = pval,
      AIC = AIC(mod)
    )
  })
})

results_clean <- results %>%
  mutate(
    mop_num = mop_to_num(mop),
    significance = if_else(!is.na(p_value) & p_value < 0.05, "Significant", "Not significant")
  ) %>%
  arrange(predictor, mop_num) %>%
  mutate(across(c(threshold, threshold_se, lower_95, upper_95, coefficient, AIC), ~round(.x, 3)),
         p_value = round(p_value, 6))

results_clean <- results_clean %>% mutate(in_main = mop %in% mops_main)
results_main  <- results_clean %>% filter(in_main)

stopifnot(all(results_main$n_events >= min_events),
          !any(results_main$mop %in% exclude_mops_default))

write_csv(results_main,
  here::here(params$tables_dir, "logistic_thresholds_by_mop_predictor.csv"))
write_csv(results_clean,
  here::here(params$tables_dir, "logistic_thresholds_by_mop_predictor_all_transects.csv"))

results_main %>% count(predictor, fit_flag)
```

# Cross-predictor comparison (reviewer comments 4, 5, 7d).
```{r}
# One row per predictor, aggregated over transects with >= 2 photos.
predictor_comparison <- results_main %>%
  dplyr::filter(fit_flag == "ok") %>%
  dplyr::mutate(ci_half_width = 1.96 * threshold_se) %>%
  dplyr::group_by(predictor) %>%
  dplyr::summarise(
    n_transects      = dplyr::n(),
    median_threshold = median(threshold, na.rm = TRUE),
    iqr_threshold    = IQR(threshold, na.rm = TRUE),
    median_ci_half   = median(ci_half_width, na.rm = TRUE),  # uncertainty metric
    median_AIC       = median(AIC, na.rm = TRUE),
    pct_significant  = 100 * mean(p_value < 0.05, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(dplyr::across(where(is.numeric), ~round(.x, 3))) %>%
  dplyr::arrange(median_ci_half)  # tightest uncertainty first

predictor_comparison
readr::write_csv(predictor_comparison,
  here::here(params$tables_dir, "predictor_comparison_summary.csv"))
```

# Fair AIC comparison: refit all six predictors on the SAME rows per
```{r}
# transect (complete cases across all six), then compare to the baseline.
preds    <- c("twl","runup","twl_daytime","runup_daytime","twl_max","twl_max_daytime")
baseline <- hp

delta_aic <- purrr::map_dfr(sort(unique(df_photos_max$mop)), function(mid) {
  d <- df_photos_max %>%
    dplyr::filter(mop == mid) %>%
    tidyr::drop_na(dplyr::all_of(c("photo", preds)))
  if (nrow(d) < 10 || dplyr::n_distinct(d$photo) < 2) return(tibble::tibble())
  aics <- vapply(preds, function(p) {
    m <- tryCatch(glm(reformulate(p, "photo"), binomial, d), error = function(e) NULL)
    if (is.null(m)) NA_real_ else AIC(m)
  }, numeric(1))
  tibble::tibble(mop = mid, predictor = preds, dAIC = aics - aics[[baseline]])
})

deltaAIC_summary <- delta_aic %>%
  dplyr::group_by(predictor) %>%
  dplyr::summarise(
    n_transects = dplyr::n(),
    median_dAIC = median(dAIC, na.rm = TRUE),   # >0 = worse than baseline
    n_better    = sum(dAIC < -2, na.rm = TRUE), # clearly better (dAIC < -2)
    n_worse     = sum(dAIC >  2, na.rm = TRUE),
    .groups = "drop"
  ) %>% dplyr::arrange(median_dAIC)

deltaAIC_summary
```

```{r}
summarise_predictors <- function(res, daic, sample_label) {
  thr <- res %>%
    filter(fit_flag == "ok") %>%
    group_by(predictor) %>%
    summarise(n_thresholds     = n(),
              median_threshold = median(threshold),
              iqr_threshold    = IQR(threshold),
              median_ci_half   = median(1.96 * threshold_se, na.rm = TRUE),
              .groups = "drop")

  cmp <- daic %>%
    filter(!is.na(dAIC)) %>%
    group_by(predictor) %>%
    summarise(n_transects = n(),
              median_dAIC = median(dAIC),
              n_worse_any = sum(dAIC > 0),
              n_better_2  = sum(dAIC < -2),
              n_worse_2   = sum(dAIC > 2),
              .groups = "drop") %>%
    mutate(p_sign = purrr::map2_dbl(n_worse_any, n_transects,
                                    ~ binom.test(.x, .y)$p.value),
           p_sign = if_else(predictor == baseline, NA_real_, p_sign))

  left_join(cmp, thr, by = "predictor") %>%
    mutate(sample = sample_label, .before = 1)
}

table_sx <- bind_rows(
  summarise_predictors(results_main,
                       filter(delta_aic, mop %in% mops_main),
                       "Main: >=2 flood photos"),
  summarise_predictors(filter(results_clean, !mop %in% exclude_mops_default),
                       filter(delta_aic, !mop %in% exclude_mops_default),
                       "Sensitivity: all transects (>=1 flood photo)")
) %>%
  arrange(factor(sample, levels = unique(sample)), match(predictor, preds_all)) %>%
  mutate(across(where(is.double), ~ signif(.x, 3)))

table_sx
readr::write_csv(table_sx, here::here(params$tables_dir, "table_SX_photo_threshold_sensitivity.csv"))
```


### Threshold CI plot
```{r}
# -----------------------------
# Settings (manuscript figure)
# -----------------------------
pname  <- hp
y_lims <- c(0, 7)

# Display name (optional; used for axis label)
predictor_display <- c(
  twl = "TWL (m)",
  runup = "Runup (m)",
  twl_daytime = "TWL-daytime (m)",
  runup_daytime = "Runup-daytime (m)",
  twl_max = "TWL-max (m)",
  twl_max_daytime = "TWL-max-daytime (m)"
)


# -----------------------------
# Build plotting dataset from results_clean
# -----------------------------
df_one <- results_main %>%
  filter(
    predictor == pname,
    fit_flag == "ok"
  ) %>%
  mutate(
    region     = mop_region(mop),
    region_col = region_colors[region],
    photo_cat  = photo_bin(n_events)
  ) %>%
  arrange(mop_num) %>%
  mutate(
    mop = factor(mop, levels = unique(mop)),
    mop_label = paste0("<span style='color:", region_col, "'>", mop, "</span>")
  )

stopifnot(nrow(df_one) > 0)

# Region label positions (centered under each region's set of MOPs)
label_positions <- df_one %>%
  filter(region %in% c("Imperial Beach", "La Jolla", "Cardiff State Beach")) %>%
  group_by(region) %>%
  summarize(x_pos = mean(as.numeric(mop)), .groups = "drop")

# Slightly above bottom axis (within y limits)
y_offset <- y_lims[1] + 0.05 * diff(y_lims)

# -----------------------------
# Plot
# -----------------------------
p_ci <- ggplot(df_one, aes(x = mop, y = threshold, color = photo_cat)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = lower_95, ymax = upper_95), width = 0.2) +
  scale_color_manual(
    values = photo_colors,
    breaks = names(photo_colors),
    drop   = FALSE
  ) +
  scale_x_discrete(labels = setNames(gsub("<[^>]+>", "", df_one$mop_label), df_one$mop)) +
  scale_y_continuous(
    limits = y_lims,
    expand = c(0, 0)
  ) +
  labs(
    x = "MOP transect",
    y = paste0("Threshold for ", predictor_display[[pname]], " (P = 0.5)"),
    color = "Total photos"
  ) +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(angle = 90, hjust = 0.5, vjust = 0.5, size = 10),
    axis.text.y  = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.title = element_text(size = 11),
    legend.text  = element_text(size = 10),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid   = element_blank()
  ) +
  annotate(
    "text",
    x = label_positions$x_pos[label_positions$region == "Imperial Beach"],
    y = y_offset,
    label = "Imperial Beach",
    color = region_colors["Imperial Beach"],
    size = 4.5
  ) +
  annotate(
    "text",
    x = label_positions$x_pos[label_positions$region == "La Jolla"],
    y = y_offset,
    label = "La Jolla",
    color = region_colors["La Jolla"],
    size = 4.5
  ) +
  annotate(
    "text",
    x = label_positions$x_pos[label_positions$region == "Cardiff State Beach"],
    y = y_offset,
    label = "Cardiff State Beach",
    color = region_colors["Cardiff State Beach"],
    size = 4.5,
    hjust = 0.7
  )

p_ci <- p_ci +
  theme(
    panel.grid.major = element_line(linewidth = 0.3, color = "grey85"),
    panel.grid.minor = element_blank()  # or element_line(linewidth = 0.2, color="grey92")
  )
p_ci
save_gg(p_ci, "f10.png", width = 10, height = 6)
```

# 30 Sept 2026: Fieller function and comparison table
```{r}
# ============================================================
# Fieller vs delta-method 95% intervals for the P = 0.5 threshold
# Predictor: headline predictor (same as Figure 9)
# ============================================================

# Fieller interval for theta = -b0/b1 from a fitted binomial GLM.
# Invert the Wald test of H0: b0 + theta*b1 = 0, i.e. solve
#   A*theta^2 + 2*B*theta + C <= 0
# where A = b1^2 - z^2*V11, B = b0*b1 - z^2*V01, C = b0^2 - z^2*V00.
# A > 0 is equivalent to the slope being significant at the chosen level;
# only then is the Fieller set a bounded interval.
fieller_threshold_ci <- function(model, level = 0.95) {
  b  <- coef(model)
  V  <- suppressWarnings(vcov(model))
  z2 <- qnorm(1 - (1 - level) / 2)^2

  na_out <- tibble(lower = NA_real_, upper = NA_real_,
                   bounded = NA, fieller_type = NA_character_)
  if (length(b) < 2 || any(!is.finite(b[1:2])) ||
      any(!is.finite(V[1:2, 1:2]))) return(na_out)

  b0 <- unname(b[1]); b1 <- unname(b[2])
  A  <- b1^2 - z2 * V[2, 2]
  B  <- b0 * b1 - z2 * V[1, 2]
  C  <- b0^2 - z2 * V[1, 1]
  disc <- B^2 - A * C

  if (A > 0 && disc >= 0) {
    tibble(lower = (-B - sqrt(disc)) / A,
           upper = (-B + sqrt(disc)) / A,
           bounded = TRUE, fieller_type = "bounded")
  } else {
    # Slope not distinguishable from zero: the confidence set is unbounded
    tibble(lower = -Inf, upper = Inf,
           bounded = FALSE, fieller_type = "unbounded")
  }
}

pname <- hp
include_single_photo <- FALSE   # TRUE adds 1-photo transects (reviewer L229 sensitivity)

delta_tbl <- results_clean %>%
  filter(predictor == pname, fit_flag == "ok",
         !mop %in% exclude_mops_default)
if (!include_single_photo) delta_tbl <- delta_tbl %>% filter(mop %in% mops_main)

delta_tbl <- delta_tbl %>%
  select(mop, n_events, total_photos, threshold, threshold_se,
         lower_95, upper_95, p_value)

# Refit each transect with the same rows used in the main loop
fieller_tbl <- purrr::map_dfr(delta_tbl$mop, function(mid) {
  d   <- df_photos_max %>%
    filter(mop == mid, !is.na(.data[[pname]]), !is.na(photo))
  fit <- fit_logit_threshold(d, pname)
  fieller_threshold_ci(fit$model) %>%
    mutate(mop = mid, threshold_refit = fit$threshold)
}) %>%
  rename(lower_fieller = lower, upper_fieller = upper)

cmp <- delta_tbl %>%
  left_join(fieller_tbl, by = "mop") %>%
  mutate(
    width_delta   = upper_95 - lower_95,
    width_fieller = upper_fieller - lower_fieller,
    width_ratio   = width_fieller / width_delta,
    lower_shift   = lower_fieller - lower_95,
    upper_shift   = upper_fieller - upper_95,
    region        = mop_region(mop),
    photo_cat     = factor(photo_bin(n_events), levels = names(photo_colors))
  )

# Sanity check: refit reproduces the stored thresholds
# (results_clean is rounded to 3 decimals, hence the tolerance)
stopifnot(all(abs(cmp$threshold - cmp$threshold_refit) < 1e-3))

# Summary by photo-count bin
cmp_summary <- cmp %>%
  group_by(photo_cat) %>%
  summarise(
    n_transects          = n(),
    n_fieller_unbounded  = sum(!bounded),
    median_width_delta   = median(width_delta),
    median_width_fieller = median(width_fieller[bounded]),
    median_width_ratio   = median(width_ratio[bounded]),
    median_lower_shift   = median(lower_shift[bounded]),
    median_upper_shift   = median(upper_shift[bounded]),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))

print(cmp_summary)
print(cmp %>% filter(!bounded) %>% select(mop, n_events, threshold, p_value))

write_csv(cmp %>% mutate(across(where(is.numeric), ~ round(.x, 3))),
          here::here(params$tables_dir, "threshold_ci_delta_vs_fieller.csv"))
```

Comparison table
```{r}
y_lims   <- c(0, 7)
y_offset <- y_lims[1] + 0.05 * diff(y_lims)

mop_order <- cmp %>% distinct(mop) %>% arrange(mop_to_num(mop)) %>% pull(mop)

lab_df <- tibble(mop = mop_order) %>%
  mutate(region = mop_region(mop),
         col    = region_colors[region],
         label  = paste0("<span style='color:", col, "'>", mop, "</span>"))

# Dodge shared by the error bars, points and clipping arrows.
# Two methods per transect, so each sits at +/- dodge_w / 4 from the transect position.
dodge_w <- 0.65

df_long <- bind_rows(
  cmp %>% transmute(mop, photo_cat, threshold, method = "Delta (Wald)",
                    lo = lower_95, hi = upper_95),
  cmp %>% transmute(mop, photo_cat, threshold, method = "Fieller",
                    lo = lower_fieller, hi = upper_fieller)
) %>%
  mutate(
    mop     = factor(mop, levels = mop_order),
    method  = factor(method, levels = c("Delta (Wald)", "Fieller")),
    lo_plot = pmax(lo, y_lims[1]),
    hi_plot = pmin(hi, y_lims[2]),
    lo_clip = lo < y_lims[1],
    hi_clip = hi > y_lims[2],
    # x position after dodging. geom_segment does not dodge xend, so the arrows
    # are placed by hand to line up with the dodged error bars.
    x_dodged = as.numeric(mop) + (as.integer(method) - 1.5) * dodge_w / 2
  )

region_lab <- lab_df %>%
  mutate(x = row_number()) %>%
  filter(region != "Other") %>%
  group_by(region) %>%
  summarise(x = mean(x), .groups = "drop")

dodge <- position_dodge(width = dodge_w)

arrow_len <- unit(0.15, "cm")   # same as f11
cap       <- 0.35               # arrow shaft length in data units (same as f11)

p_cmp <- ggplot(df_long,
                aes(x = mop, y = threshold, color = photo_cat,
                    shape = method, linetype = method, group = method)) +
  geom_errorbar(aes(ymin = lo_plot, ymax = hi_plot),
                width = 0.35, position = dodge) +
  geom_point(size = 2, position = dodge) +
  # Arrows point outward where the interval runs past the top of the axis ...
  geom_segment(
    data = filter(df_long, hi_clip),
    aes(x = x_dodged, xend = x_dodged,
        y = y_lims[2] - cap, yend = y_lims[2], color = photo_cat),
    inherit.aes = FALSE, show.legend = FALSE,
    arrow = arrow(length = arrow_len)
  ) +
  # ... or past the bottom of the axis (unbounded Fieller intervals get both)
  geom_segment(
    data = filter(df_long, lo_clip),
    aes(x = x_dodged, xend = x_dodged,
        y = y_lims[1] + cap, yend = y_lims[1], color = photo_cat),
    inherit.aes = FALSE, show.legend = FALSE,
    arrow = arrow(length = arrow_len)
  ) +
  geom_text(data = region_lab, aes(x = x, y = y_offset, label = region),
            inherit.aes = FALSE, size = 4.5,
            color = region_colors[region_lab$region]) +
  scale_color_manual(values = photo_colors, breaks = names(photo_colors),
                     drop = FALSE) +
  scale_shape_manual(values = c("Delta (Wald)" = 16, "Fieller" = 15)) +
  scale_linetype_manual(values = c("Delta (Wald)" = "solid", "Fieller" = "dashed")) +
  scale_x_discrete(labels = setNames(gsub("<[^>]+>", "", lab_df$label), lab_df$mop)) +
  scale_y_continuous(limits = y_lims, expand = c(0, 0)) +
  labs(x = "MOP transect",
       y = paste0("Threshold for ", hp_name, " (P = 0.5)"),
       color = "Total photos", shape = "Interval", linetype = "Interval") +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(angle = 90, hjust = 0.5, vjust = 0.5, size = 10),
    axis.text.y  = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.title = element_text(size = 11),
    legend.text  = element_text(size = 10),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid.major = element_line(linewidth = 0.3, color = "grey85"),
    panel.grid.minor = element_blank()
  )

p_cmp
save_gg(p_cmp, "f_delta_vs_fieller.png", width = 10, height = 6)
```

Detailed analysis
```{r}
results_main %>%
  filter(predictor == hp) %>%
  count(fit_flag)

results_clean %>%
  distinct(mop, total_photos, n_events) %>%
  filter(total_photos != n_events)
```
Following up to see if any of the single photo MOP transects have a flag
```{r}
# 1. Flag counts by predictor, all transects in results_clean (D0841 excluded)
results_clean %>%
  filter(!mop %in% exclude_mops_default) %>%
  count(predictor, fit_flag) %>%
  pivot_wider(names_from = fit_flag, values_from = n, values_fill = 0)

# 2. Every non-ok fit, any predictor
results_clean %>%
  filter(fit_flag != "ok") %>%
  select(mop, predictor, total_photos, n_events, fit_flag, slope, p_value) %>%
  arrange(predictor, mop)

# 3. TWL-max-daytime only, split by one photo vs two or more
results_clean %>%
  filter(predictor == hp, !mop %in% exclude_mops_default) %>%
  mutate(photo_group = if_else(n_events == 1, "1 photo", "2+ photos")) %>%
  count(photo_group, fit_flag)
```



```{r}
# ============================
# Summary stats for f10
#
# Output thresholds for TWL and TWL-max and TWL-max-daytime
# with 95% CI half-width column
# ============================

predictors_to_output <- unique(c("twl", "twl_max", "twl_max_daytime", hp))

df_thresholds_two <- results_main %>%
  filter(
    predictor %in% predictors_to_output
  ) %>%
  mutate(
    ci_half_width = 1.96 * threshold_se  # 95% CI half-width
  ) %>%
  select(
    mop,
    predictor,
    total_photos,
    threshold,
    threshold_se,
    ci_half_width,
    lower_95,
    upper_95
  ) %>%
  arrange(predictor, mop)

print(df_thresholds_two)
```

Mean threshold +/- SD
```{r}
preds <- c("twl", "twl_max", "twl_daytime",
           "twl_max_daytime", "runup", "runup_daytime")

summarise_thresholds <- function(res) {
  res %>%
    filter(predictor %in% preds, fit_flag == "ok") %>%
    group_by(predictor) %>%
    summarise(
      n_transects      = n_distinct(mop),
      threshold_mean   = mean(threshold),
      threshold_sd     = sd(threshold),
      threshold_median = median(threshold),
      q1               = quantile(threshold, 0.25),
      q3               = quantile(threshold, 0.75),
      iqr              = IQR(threshold),
      .groups = "drop"
    ) %>%
    arrange(match(predictor, preds))
}

summary_main <- summarise_thresholds(results_main)
summary_main
```
# Filter to statistically significant models
```{r}

alpha <- 0.05

summary_main_sig <- summarise_thresholds(filter(results_main, p_value < alpha))
summary_main_sig
```


# ============================================================
# Bayesian Analysis: Transect-specific TWL thresholds (P=0.5)
# Outcome: photo (0/1)
# Predictor: headline predictor (centered at params$twl_threshold_for_summary)
# ============================================================


```{r}

# ---- Convert draws to threshold draws ----
# If model is: logit(p) = a + b * x_centered
# then p=0.5 occurs at a + b*x = 0  -> x* = -a/b
# uncenter: TWL* = center_value + x*
threshold_draws_from_fit <- function(fit, center_value) {
  d <- posterior::as_draws_df(fit)

  # brms naming: b_Intercept and b_<predictor>
  if (!("b_Intercept" %in% names(d))) stop("Could not find b_Intercept in posterior draws.")

  slope_name <- "b_x_centered"
  if (!(slope_name %in% names(d))) stop("Missing slope draws: ", slope_name)
  if (length(slope_name) != 1) {
    stop("Could not uniquely identify slope draws for x_centered.")
  }

  a <- d[["b_Intercept"]]
  b <- d[[slope_name]]

  # Guard against division by ~0 slopes
  eps <- 1e-6
  thr_centered <- -a / b
  thr_uncentered <- center_value + thr_centered
  thr_uncentered[is.finite(thr_uncentered) & abs(b) > eps]
}
```

```{r}
# Use df_photos_max from earlier pipeline
# Require predictor + outcome
df_bayes <- df_photos_max %>%
  filter(!is.na(photo), !is.na(.data[[hp]])) %>%
  mutate(
    x_centered = .data[[hp]] - params$twl_threshold_for_summary
  )

# Photo counts per MOP (flood photos only)
df_photo_counts <- df_bayes %>%
  group_by(mop) %>%
  summarise(
    n_flood_photos = sum(photo == 1, na.rm = TRUE),
    n_total_days   = n(),
    .groups = "drop"
  )

# Use the same transect sample as the frequentist analysis (defined in Data preparation):
# >= min_events flood photos on complete-case rows, excluding data-quality drops (e.g., D0841).
mops_keep <- mops_main

# One-time check: which transects did the old rule select differently?
mops_keep_old <- df_photo_counts %>%
  filter(n_flood_photos >= min_events, !mop %in% exclude_mops_default) %>%
  pull(mop)
setdiff(mops_keep_old, mops_main)
setdiff(mops_main, mops_keep_old)

length(mops_keep)
```


```{r}
# ------------------------------------------------------------
# Model spec
# ------------------------------------------------------------
# Weakly informative priors (WEAKER), interpretable with centering at 3.4 m:
# - Intercept ~ Normal(0, 2.5): log-odds at TWL=3.4 (broad)
# - Slope     ~ Normal(0, 2.5): allows positive/negative slopes; less prior push
priors <- c(
  prior(normal(0, 2.5), class = "Intercept"),
  prior(normal(0, 2.5), class = "b")
)

# ------------------------------------------------------------
# Compile ONCE using one MOP with adequate data/variability
# (pick the first MOP that actually has enough rows + both classes)
# ------------------------------------------------------------
example_mop <- mops_keep[
  which(map_lgl(mops_keep, function(m) {
    dd <- df_bayes %>%
      filter(mop == m) %>%
      filter(!is.na(photo), !is.na(x_centered))
    nrow(dd) >= 10 && n_distinct(dd$photo) >= 2
  }))[1]
]

if (is.na(example_mop)) {
  stop("No MOP in mops_keep has >=10 rows and both photo classes for model compilation.")
}

example_data <- df_bayes %>%
  filter(mop == example_mop) %>%
  filter(!is.na(photo), !is.na(x_centered))

base_fit <- brm(
  formula = photo ~ x_centered,
  data    = example_data,
  family  = bernoulli(link = "logit"),
  prior   = priors,
  chains  = 2,
  iter    = 1000,    # quick compile fit
  warmup  = 500,
  seed    = 123,
  refresh = 0
)

# ------------------------------------------------------------
# Loop over transects, reusing compiled Stan program
# ------------------------------------------------------------
bayes_results <- tibble(
  mop              = character(),
  n_flood_photos   = integer(),
  threshold_median = double(),
  threshold_lower  = double(),
  threshold_upper  = double(),
  max_rhat         = double(),
  min_ess_bulk     = double(),
  min_ess_tail     = double(),
  n_divergent      = integer()
)

all_draws_ge2 <- list()
all_draws_ge8 <- list()

for (m in mops_keep) {

  d_m <- df_bayes %>%
    filter(mop == m) %>%
    filter(!is.na(photo), !is.na(x_centered))

  if (nrow(d_m) < 10 || n_distinct(d_m$photo) < 2) next

  fit_m <- update(
    base_fit,
    newdata   = d_m,
    recompile = FALSE,
    chains    = 4,
    iter      = 2000,
    warmup    = 1000,
    seed      = 123,
    refresh   = 0,
    control   = list(adapt_delta = 0.95, max_treedepth = 12)
  )
  
  # --- Convergence diagnostics: substantiate the Methods claim (per-fit) ---
  diag_tbl <- tryCatch(
    posterior::summarise_draws(
      posterior::as_draws_df(fit_m), "rhat", "ess_bulk", "ess_tail"
    ),
    error = function(e) NULL
  )
  if (is.null(diag_tbl)) {
    max_rhat <- NA_real_; min_ess_bulk <- NA_real_; min_ess_tail <- NA_real_
  } else {
    max_rhat     <- max(diag_tbl$rhat,     na.rm = TRUE)
    min_ess_bulk <- min(diag_tbl$ess_bulk, na.rm = TRUE)
    min_ess_tail <- min(diag_tbl$ess_tail, na.rm = TRUE)
  }
  n_div <- tryCatch({
    np <- brms::nuts_params(fit_m)
    sum(np$Value[np$Parameter == "divergent__"])
  }, error = function(e) NA_real_)

  thr_draws <- threshold_draws_from_fit(
    fit_m,
    center_value = params$twl_threshold_for_summary
  )

  qs <- quantile(thr_draws, probs = c(0.025, 0.5, 0.975), na.rm = TRUE)

  n_flood <- df_photo_counts %>%
    filter(mop == m) %>%
    pull(n_flood_photos)

  if (length(n_flood) != 1 || is.na(n_flood)) next

  bayes_results <- bayes_results %>%
    add_row(
      mop              = m,
      n_flood_photos   = as.integer(n_flood),
      threshold_median = as.numeric(qs[[2]]),
      threshold_lower  = as.numeric(qs[[1]]),
      threshold_upper  = as.numeric(qs[[3]]),
      max_rhat         = max_rhat,
      min_ess_bulk     = min_ess_bulk,
      min_ess_tail     = min_ess_tail,
      n_divergent      = as.integer(n_div)
    )

  if (n_flood >= 2) all_draws_ge2[[m]] <- thr_draws
  if (n_flood >= 8) all_draws_ge8[[m]] <- thr_draws

  if (isTRUE(params$dev_mode)) message("Bayes processed: ", m, " (n_flood_photos=", n_flood, ")")
}

bayes_results <- bayes_results %>%
  mutate(
    mop_num    = mop_to_num(mop),
    region     = mop_region(mop),
    region_col = region_colors[region],
    photo_cat  = photo_bin(n_flood_photos)
  ) %>%
  arrange(mop_num) %>%
  mutate(
    mop = factor(mop, levels = unique(mop)),
    mop_label = paste0("<span style='color:", region_col, "'>", as.character(mop), "</span>")
  )

bayes_results %>% glimpse()

out_bayes <- here::here(params$tables_dir, "bayes_thresholds_by_mop.csv")
readr::write_csv(bayes_results, out_bayes)
if (isTRUE(params$dev_mode)) message("Saved: ", out_bayes)
```
```{r bayes-convergence-summary}
# One-line convergence summary for Methods/Results wording.
convergence_summary <- bayes_results %>%
  dplyr::summarise(
    n_models         = dplyr::n(),
    max_rhat_overall = max(max_rhat, na.rm = TRUE),
    min_ess_bulk     = min(min_ess_bulk, na.rm = TRUE),
    min_ess_tail     = min(min_ess_tail, na.rm = TRUE),
    total_divergent  = sum(n_divergent, na.rm = TRUE),
    n_rhat_gt_1_01   = sum(max_rhat > 1.01, na.rm = TRUE)
  )
print(convergence_summary)
```


```{r}
# ============================================================
# Figure 11: Bayesian thresholds (median + 95% credible interval)
# Colored by flood-photo count bins; x tick labels colored by region.
# Shows truncated intervals WITHIN plot bounds + arrows when clipped.
# ============================================================

# Optional exclusions (match your earlier choices if needed)
df_f11 <- bayes_results %>%
  filter(!mop %in% exclude_mops_default)

# y-limits for the figure (tune if you want)
y_lims <- c(0, 7)
y_offset <- y_lims[1] + 0.05 * diff(y_lims)

# ---- Truncate CI to plotting window + flags for clipping ----
df_f11 <- df_f11 %>%
  mutate(
    lower_plot = pmax(threshold_lower, y_lims[1]),
    upper_plot = pmin(threshold_upper, y_lims[2]),
    lower_trunc = threshold_lower < y_lims[1],
    upper_trunc = threshold_upper > y_lims[2]
  )

# Region label positions (centered under each region's MOP group)
label_positions <- df_f11 %>%
  filter(region %in% c("Imperial Beach", "La Jolla", "Cardiff State Beach")) %>%
  group_by(region) %>%
  summarize(x_pos = mean(as.numeric(mop)), .groups = "drop")

# Arrow size and small offset (in data units) so arrows are visible
arrow_len <- unit(0.15, "cm")
cap <- 0.35  # length of the arrow shaft in data units

p_f11 <- ggplot(df_f11, aes(x = mop, y = threshold_median, color = photo_cat)) +
  geom_point(size = 2) +

  # Draw truncated error bars (always within y_lims)
  geom_errorbar(aes(ymin = lower_plot, ymax = upper_plot), width = 0.2) +

  # Add bottom arrows where lower bound is below plot
  geom_segment(
    data = df_f11 %>% filter(lower_trunc),
    aes(x = mop, xend = mop, y = y_lims[1] + cap, yend = y_lims[1]),
    inherit.aes = FALSE,
    arrow = arrow(length = arrow_len)
  ) +

  # Add top arrows where upper bound is above plot
  geom_segment(
    data = df_f11 %>% filter(upper_trunc),
    aes(x = mop, xend = mop, y = y_lims[2] - cap, yend = y_lims[2]),
    inherit.aes = FALSE,
    arrow = arrow(length = arrow_len)
  ) +

  scale_color_manual(values = photo_colors, breaks = names(photo_colors), drop = FALSE) +
  scale_x_discrete(labels = setNames(gsub("<[^>]+>", "", df_f11$mop_label), df_f11$mop)) +
  scale_y_continuous(limits = y_lims, expand = c(0, 0)) +
  labs(
    x = "MOP transect",
    y = paste0(hp_name, " threshold (m) for P(photo)=0.5"),
    color = "Total photos"
  ) +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(angle = 90, hjust = 0.5, vjust = 0.5, size = 10),
    axis.text.y  = element_text(size = 10),
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    legend.title = element_text(size = 11),
    legend.text  = element_text(size = 10),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid   = element_blank(),
    panel.grid.major = element_line(linewidth = 0.3, color = "grey85"),
    panel.grid.minor = element_blank()
  ) +
  annotate(
    "text",
    x = label_positions$x_pos[label_positions$region == "Imperial Beach"],
    y = y_offset,
    label = "Imperial Beach",
    color = region_colors["Imperial Beach"],
    size = 4.5
  ) +
  annotate(
    "text",
    x = label_positions$x_pos[label_positions$region == "La Jolla"],
    y = y_offset,
    label = "La Jolla",
    color = region_colors["La Jolla"],
    size = 4.5
  ) +
  annotate(
    "text",
    x = label_positions$x_pos[label_positions$region == "Cardiff State Beach"],
    y = y_offset,
    label = "Cardiff State Beach",
    color = region_colors["Cardiff State Beach"],
    size = 4.5,
    hjust = 0.7
  )

p_f11
save_gg(p_f11 + labs(title = NULL), "f11.png", width = 10, height = 6)
```

```{r}
# ============================================================
# Manuscript-ready summary numbers
# (1) D0038
# (2) "Countywide" pooled posterior across transects with >=2 flood photos
# (3) Pooled posterior across transects with >=8 flood photos
# ============================================================

# ---- (1) D0038 ----
d0038 <- bayes_results %>%
  filter(as.character(mop) == "D0038") %>%
  slice(1)

d0038

# ---- Helper: pooled posterior across transects (equal-weight by transect) ----
pooled_summary <- function(draws_list) {
  pooled <- unlist(draws_list, use.names = FALSE)
  qs <- quantile(pooled, probs = c(0.025, 0.5, 0.975), na.rm = TRUE)
  tibble(
    pooled_median = as.numeric(qs[[2]]),
    pooled_lower  = as.numeric(qs[[1]]),
    pooled_upper  = as.numeric(qs[[3]])
  )
}

pool_ge2 <- pooled_summary(all_draws_ge2)
pool_ge8 <- pooled_summary(all_draws_ge8)

pool_ge2
pool_ge8

# ---- Print manuscript-ready sentences (rounded) ----
fmt <- function(x) sprintf("%.2f", x)

cat(
  "Bayesian D0038 threshold: ",
  fmt(d0038$threshold_median),
  " m [",
  fmt(d0038$threshold_lower),
  ", ",
  fmt(d0038$threshold_upper),
  "].\n",
  sep = ""
)

# Summary across the main transects (>= 2 flood photos): median and range of
# the transect-level posterior medians. Stacking draws across transects is not
# reported for this set because sparsely documented transects dominate it.
main_summary <- bayes_results %>%
  summarise(n = n(),
            median_of_medians = median(threshold_median),
            min_median = min(threshold_median),
            max_median = max(threshold_median))
cat(sprintf(
  "Transect-level posterior medians (n = %d, >=2 flood photos): median %.2f m, range %.2f to %.2f m.\n",
  main_summary$n, main_summary$median_of_medians,
  main_summary$min_median, main_summary$max_median))

cat(
  "NOT REPORTED - stacked draws, >=2 flood photos: ",
  fmt(pool_ge2$pooled_median),
  " m [",
  fmt(pool_ge2$pooled_lower),
  ", ",
  fmt(pool_ge2$pooled_upper),
  "].\n",
  sep = ""
)

cat(
  "Pooled (>=8 flood photos) threshold: ",
  fmt(pool_ge8$pooled_median),
  " m [",
  fmt(pool_ge8$pooled_lower),
  ", ",
  fmt(pool_ge8$pooled_upper),
  "].\n",
  sep = ""
)
```



# Density plots
```{r}
# -----------------------------
# Settings for f12 (IB only)
# -----------------------------
pname <- hp

# -----------------------------
# Prep plotting data (IB ONLY)
# -----------------------------
df_f12 <- df_photos_max %>%
  filter(mop >= params$ib_mop_min, mop <= params$ib_mop_max) %>%  # IB filter
  filter(mop %in% mops_main) %>%                                  # shared sample (>= min_events, exclusions applied)
  filter(!is.na(photo), !is.na(.data[[pname]]))

stopifnot(nrow(df_f12) > 0)

# Facet labels: drop D00 prefix for compactness
df_f12 <- df_f12 %>%
  mutate(
    mop_short = str_replace(mop, "^D00", ""),
    mop_num   = as.numeric(gsub("^D", "", mop))
  ) %>%
  arrange(mop_num) %>%
  mutate(mop_short = factor(mop_short, levels = unique(mop_short)))

# -----------------------------
# Logistic thresholds per MOP
# -----------------------------
thr_f12 <- df_f12 %>%
  group_by(mop, mop_short) %>%
  summarize(
    threshold = tryCatch(
      fit_logit_threshold(cur_data(), pname)$threshold,
      error = function(e) NA_real_
    ),
    .groups = "drop"
  ) %>%
  filter(!is.na(threshold))

# Balanced documentation threshold per transect (fitted probability = overall photo frequency)
balanced_f12 <- df_f12 %>%
  group_by(mop, mop_short) %>%
  group_modify(~ {
    f <- fit_logit_threshold(.x, pname)
    tibble(balanced_thr = f$threshold + qlogis(mean(.x$photo)) / f$slope)
  }) %>%
  ungroup() %>%
  filter(!is.na(balanced_thr))

f12_check <- thr_f12 %>%
  left_join(results_main %>% filter(predictor == pname) %>% select(mop, threshold),
            by = "mop", suffix = c("_f12", "_f10")) %>%
  mutate(diff = threshold_f12 - threshold_f10) %>%
  summarise(max_abs_diff = max(abs(diff), na.rm = TRUE)) %>%
  pull(max_abs_diff)

tol <- 6e-4  # results_main thresholds are rounded to 3 decimals
cat(sprintf(
  "Fig 12 vs Fig 10 thresholds: max abs difference = %.2e (expected 0 or < %.0e) -> %s\n",
  f12_check, tol, if (f12_check < tol) "PASS" else "CHECK: figures fit on different rows"
))

# -----------------------------
# Plot
# -----------------------------
p_f12 <- ggplot(df_f12, aes(x = .data[[pname]], fill = factor(photo))) +
  geom_density(alpha = 0.5, position = "identity") +
  scale_fill_manual(
    values = c("0" = "blue", "1" = "red"),
    name = "Photo",
    labels = c("0" = "No Photo", "1" = "Photo")
  ) +
  facet_wrap(~ mop_short, ncol = 1, strip.position = "left") +
  geom_vline(
    data = thr_f12,
    aes(xintercept = threshold),
    color = "black",
    linetype = "dashed",
    linewidth = 0.6,
    inherit.aes = FALSE
  ) +
  geom_vline(
    data = balanced_f12,
    aes(xintercept = balanced_thr),
    color = "darkorange",
    linetype = "longdash",
    linewidth = 0.6,
    inherit.aes = FALSE
  ) +
  labs(
    x = hp_label,
    y = "Density",
    title = NULL
  ) +
  theme_minimal() +
  theme(
    panel.spacing = unit(0.08, "lines"),
    strip.text.y  = element_text(size = 10),
    axis.text.y   = element_blank(),
    axis.ticks.y  = element_blank(),
    axis.text.x   = element_text(size = 10),
    axis.title.x  = element_text(size = 12, margin = margin(t = 10)),
    axis.title.y  = element_text(size = 12, margin = margin(r = 10)),
    panel.border  = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid    = element_blank()
  )

p_f12

# -----------------------------
# Save with height scaled to # IB facets
# -----------------------------
n_facets <- n_distinct(df_f12$mop_short)
h_in <- max(6, 0.55 * n_facets)   # IB-only usually looks good with ~0.55 per facet

save_gg(p_f12, "f12.png", width = 9, height = h_in, dpi = 300)
```

# Atmospheric River Analysis

```{r}
# ============================================================
# Atmospheric River Analysis (Manuscript Figure f13)
# Question: Is a photo more likely to occur within a window
#          AFTER an AR made landfall?
#
# Operationalization (daily):
#   AR_after_k = 1 if an AR (ARsr=1) occurred on the same day
#               OR within the previous k days (lookback window).
#   Interpretation: on a photo day, was there an AR within the last k days (inclusive)?
#   This is equivalent to: "photo day falls within k days after an AR day."
#
# Model (one per k): photo ~ AR_after_k + (1 | mop)
# Output: f13.png
# ============================================================
```



```{r f13-ar-after-window, eval=has_ar}

# -----------------------------
# Read AR catalog + define ARsr
# -----------------------------
ar_path <- here::here(params$data_dir, params$ar_catalog_xlsx)
assert_file_exists(ar_path)

ar <- readxl::read_excel(ar_path, sheet = params$ar_catalog_sheet) %>%
  mutate(
    Date = as.Date(Date),
    ARsr = ifelse(!is.na(StormRank) & StormRank > 0, 1L, 0L)
  ) %>%
  select(Date, ARsr) %>%
  group_by(Date) %>%
  summarise(ARsr = max(ARsr, na.rm = TRUE), .groups = "drop")

# -----------------------------
# Merge AR onto TWL/photo daily rows
# -----------------------------
df_ar <- df_photos_max %>%
  mutate(date = as.Date(date)) %>%
  select(mop, date, photo, twl, twl_daytime, twl_max, twl_max_daytime) %>%
  distinct(mop, date, .keep_all = TRUE) %>%   # safety
  left_join(ar, by = c("date" = "Date")) %>%
  mutate(
    photo = ifelse(is.na(photo), 0L, as.integer(photo)),
    ARsr  = ifelse(is.na(ARsr),  0L, as.integer(ARsr)),
    mop   = factor(mop)
  ) %>%
  arrange(mop, date)

# -----------------------------
# Build cumulative post-AR windows:
# AR_after0 = AR on same day
# AR_after3 = AR on same day OR within previous 3 days
# ... up to K days
# -----------------------------
K <- params$ar_window_max_lag_days  # e.g., 7

df_ar <- df_ar %>%
  group_by(mop) %>%
  arrange(date, .by_group = TRUE) %>%
  group_modify(~{
    d <- .x
    x <- as.integer(d$ARsr)  # ensure plain integer vector

    for (k in 0:K) {
      # OR across lagged versions 0..k (all length nrow(d), even when nrow(d)==1)
      any_in_window <- Reduce(
        `|`,
        lapply(0:k, function(j) dplyr::lag(x, n = j, default = 0L) == 1L)
      )

      d[[paste0("AR_after", k)]] <- as.integer(any_in_window)
    }

    d
  }) %>%
  ungroup()

stopifnot(all(paste0("AR_after", 0:K) %in% names(df_ar)))
stopifnot(all(unlist(df_ar[paste0("AR_after", 0:K)]) %in% c(0L, 1L)))

# -----------------------------
# Fit GLMM for each window length
# -----------------------------
fit_one <- function(k) {
  pred <- paste0("AR_after", k)

  # If no variation, skip cleanly
  if (length(unique(df_ar[[pred]])) < 2) {
    return(tibble(
      window_days = k, estimate = NA_real_, std_error = NA_real_,
      odds_ratio = NA_real_, ci_lower = NA_real_, ci_upper = NA_real_,
      p_value = NA_real_, AIC = NA_real_
    ))
  }

  fml  <- as.formula(paste0("photo ~ ", pred, " + (1 | mop)"))
  mod <- glmer(
    fml, data = df_ar, family = binomial(link = "logit"),
    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
  )

  co <- summary(mod)$coefficients
  if (!(pred %in% rownames(co))) {
    return(tibble(
      window_days = k, estimate = NA_real_, std_error = NA_real_,
      odds_ratio = NA_real_, ci_lower = NA_real_, ci_upper = NA_real_,
      p_value = NA_real_, AIC = AIC(mod)
    ))
  }

  est <- co[pred, "Estimate"]
  se  <- co[pred, "Std. Error"]

  tibble(
    window_days = k,
    estimate    = est,
    std_error   = se,
    odds_ratio  = exp(est),
    ci_lower    = exp(est - 1.96 * se),
    ci_upper    = exp(est + 1.96 * se),
    p_value     = 2 * (1 - pnorm(abs(est / se))),
    AIC         = AIC(mod)
  )
}

ar_glmm_results <- bind_rows(lapply(0:K, fit_one)) %>%
  mutate(
    label = factor(
      window_days,
      levels = 0:K,
      labels = c("0 days post-AR", paste0("0–", 1:K, " days post-AR"))
    )
  )

# -----------------------------
# Plot + save (f13)
# -----------------------------
p_f13 <- ggplot(ar_glmm_results, aes(x = label, y = odds_ratio)) +
  geom_col(fill = "#1f77b4") +   # Blue bars
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    width = 0.2,
    color = "black"             # Black error bars
  ) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed",
    color = "red",              # Red reference line
    linewidth = 0.8
  ) +
  labs(
    x = "Post-AR window (0–k days after AR; inclusive)",
    y = "Odds ratio for Photo = 1 (GLMM; 95% CI)",
    title = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(angle = 45, hjust = 1, vjust = 1, size = 11),
    axis.text.y  = element_text(size = 11),
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid   = element_blank(),
    panel.grid.major.y = element_line(linewidth = 0.3, color = "grey85")
  )

p_f13
save_gg(p_f13, "f13.png", width = 10, height = 6, dpi = 300)
```
Linear model
```{r, eval=has_ar}
df_twl_lm <- df_ar %>%
  group_by(mop) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(AR_lag1 = lag(ARsr, 1, default = 0L)) %>%
  ungroup()

lm_fit <- lm(twl ~ AR_lag1, data = df_twl_lm)

summary(lm_fit)

```
Linear mixed-effects model
```{r, eval=has_ar}
lmm_fit <- lmer(twl ~ AR_lag1 + (1 | mop), data = df_twl_lm)
summary(lmm_fit)
```
Mean TWL on non-AR days ≈ 1.78 m
Within one dat post-AR, TWL is on average 0.13 meters higher (≈ 13 cm) than on non-AR days.
Statistical significance: t = 26.67 → extremely large
Within a one day window post-AR there was a statistically significant but modest increase in TWL (~0.13 m).
AR incidence explains only a small fraction of TWL variability relative to tides, waves, and surge.

```{r, eval=has_ar}
photo_days <- df_ar %>%
  group_by(date) %>%
  summarize(photo_day = any(photo == 1)) %>%
  filter(photo_day) %>%
  pull(date)

ar_days <- df_ar %>%
  group_by(date) %>%
  summarize(ar_day = any(ARsr == 1)) %>%
  filter(ar_day) %>%
  pull(date)

length(intersect(photo_days, ar_days))
```
MOP transect D0038 analysis
```{r, eval=has_ar}
df_d0038 <- df_ar %>%
  filter(mop == "D0038")

# Function to fit model for one window
fit_window_glm <- function(k) {

  pred <- paste0("AR_after", k)
  fml  <- as.formula(paste0("photo ~ ", pred))

  mod <- glm(
    fml,
    family = binomial(link = "logit"),
    data   = df_d0038
  )

  co <- summary(mod)$coefficients

  est <- co[pred, "Estimate"]
  se  <- co[pred, "Std. Error"]
  z   <- co[pred, "z value"]
  p   <- co[pred, "Pr(>|z|)"]

  tibble(
    window_days = k,
    estimate    = est,
    std_error   = se,
    odds_ratio  = exp(est),
    ci_lower    = exp(est - 1.96 * se),
    ci_upper    = exp(est + 1.96 * se),
    z_value     = z,
    p_value     = p,
    AIC         = AIC(mod)
  )
}

# run for 0-7 days
d0038_window_results <- bind_rows(
  lapply(0:7, fit_window_glm)
)

d0038_window_results

```
```{r, eval=has_ar}
df_d0038 <- df_ar %>% filter(mop == "D0038")

model_after3 <- glm(photo ~ AR_after3, family = binomial, data = df_d0038)
summary(model_after3)

model_after3 <- glm(photo ~ AR_after3,
                    family = binomial,
                    data = df_d0038)

summary(model_after3)
```


# ENSO Analysis
```{r, eval=has_oni}
# ============================================================
# ENSO Analysis (ONI)  -> outputs f14.png, f15.png
# ============================================================

# -----------------------------
# Read ONI netCDF (monthly)
# -----------------------------
oni_path <- here::here(params$data_dir, params$oni_nc)
assert_file_exists(oni_path)

nc <- nc_open(oni_path)
oni_time   <- ncvar_get(nc, "time")
oni_value  <- ncvar_get(nc, "value")
time_units <- ncatt_get(nc, "time", "units")$value
nc_close(nc)

# Helper: convert netCDF time units -> Date vector
nc_time_to_date <- function(t, units_str) {
  # Examples we handle:
  # "days since 1948-1-1 00:00:0.0"
  # "days since 1950-01-01"
  # "months since 1950-01-01"
  units_str <- as.character(units_str)

  if (str_detect(units_str, "^days since")) {
    origin <- str_trim(str_remove(units_str, "^days since"))
    origin <- str_extract(origin, "^\\d{4}-?\\d{1,2}-?\\d{1,2}")
    origin <- as.Date(origin)
    return(origin + days(t))
  }

  if (str_detect(units_str, "^months since")) {
    origin <- str_trim(str_remove(units_str, "^months since"))
    origin <- str_extract(origin, "^\\d{4}-?\\d{1,2}-?\\d{1,2}")
    origin <- as.Date(origin)
    # add months safely
    return(origin %m+% months(t))
  }

  stop("Unsupported time units in oni.nc: ", units_str)
}

oni_dates <- nc_time_to_date(oni_time, time_units)

oni_monthly <- tibble(
  Time  = as.Date(oni_dates),
  Value = as.numeric(oni_value)
) %>%
  arrange(Time)

# ------------------------------------------------------------
# NOAA/CPC canonical ONI event definition:
# El Niño if ONI >= +0.5 for at least 5 consecutive overlapping seasons
# La Niña if ONI <= -0.5 for at least 5 consecutive overlapping seasons
# Otherwise Neutral
# ------------------------------------------------------------

label_enso_noaa_cpc <- function(oni, thr = 0.5, min_run = 5) {
  n <- length(oni)
  phase <- rep("Neutral", n)

  el <- !is.na(oni) & oni >=  thr
  la <- !is.na(oni) & oni <= -thr

  # Find runs of TRUE of length >= min_run
  mark_runs <- function(flag, label) {
    r <- rle(flag)
    ends <- cumsum(r$lengths)
    starts <- ends - r$lengths + 1
    ok <- which(r$values & r$lengths >= min_run)
    if (length(ok) == 0) return(rep(FALSE, n))
    idx <- unlist(Map(seq, starts[ok], ends[ok]))
    out <- rep(FALSE, n)
    out[idx] <- TRUE
    out
  }

  el_idx <- mark_runs(el, "El Niño")
  la_idx <- mark_runs(la, "La Niña")

  phase[el_idx] <- "El Niño"
  phase[la_idx] <- "La Niña"

  # If something is somehow both (shouldn't happen), prioritize the larger magnitude
  both <- el_idx & la_idx
  if (any(both)) {
    phase[both] <- ifelse(oni[both] >= 0, "El Niño", "La Niña")
  }

  phase
}

oni_monthly <- oni_monthly %>%
  arrange(Time) %>%
  mutate(
    enso_phase = label_enso_noaa_cpc(Value, thr = 0.5, min_run = 5),
    enso_phase = factor(enso_phase, levels = c("El Niño", "La Niña", "Neutral"))
  )

# -----------------------------
# Expand monthly ONI -> daily (repeat value/phase each day in month)
# -----------------------------
oni_daily <- oni_monthly %>%
  mutate(
    Date = map(Time, ~ seq(.x, .x + days(days_in_month(.x) - 1), by = "day"))
  ) %>%
  unnest(Date) %>%
  select(Date, oni = Value, enso_phase)

# -----------------------------
# Join ENSO onto your TWL daily dataset
# -----------------------------
# Backbone TWL (daily, 2000–2023)
twl_path <- here::here(params$transects_rds)
assert_file_exists(twl_path)

twl_23 <- readRDS(twl_path) %>%
  mutate(date = as.Date(date)) %>%
  distinct(mop, date, .keep_all = TRUE)  # IMPORTANT: one row per mop-date

# Join ONI to TWL backbone
twl_enso <- twl_23 %>%
  left_join(oni_daily, by = c("date" = "Date"))
```

```{r, eval=has_oni}
photo_overlay <- df_photos_max %>%
  mutate(date = as.Date(date)) %>%
  # keep only what you need to overlay onto the 2000–2023 backbone
  select(
    mop, date,
    photo,
    twl_max, twl_max_daytime,
    twl_1_day_before, runup_1_day_before,
    twl_daytime, runup, runup_daytime
  ) %>%
  # ensure photo is 0/1 and de-dupe mop-date (keep photo==1 if duplicates)
  mutate(photo = ifelse(is.na(photo), 0L, as.integer(photo))) %>%
  group_by(mop, date) %>%
  arrange(desc(photo)) %>%
  slice(1) %>%
  ungroup()

# Join photo overlay onto the backbone-with-ENSO
twl_enso_photos <- twl_enso %>%
  left_join(photo_overlay, by = c("mop","date"), suffix = c("", "_photo")) %>%
  mutate(photo = ifelse(is.na(photo), 0L, as.integer(photo)))
```

```{r, eval=has_oni}
# ---------------------------------------------------------------------------
# Shared settings for the ENSO figures
# ---------------------------------------------------------------------------
# ENSO figures use the daily (24-hour) maximum TWL, as in the submitted manuscript.
enso_var <- "twl"
enso_label <- unname(predictor_labels[enso_var])

# ---------------------------------------------------------------------------
# f14: D0044 time series with ENSO stripe, full record
# Changes: end date from the data (keeps 2024 photos); y-limits from the data
# with the stripe drawn above the highest peak; gap diagnostic printed.
# ---------------------------------------------------------------------------
f14_mop <- "D0044"

df_f14 <- twl_enso_photos %>%
  filter(mop == f14_mop, date >= as.Date("2000-01-01")) %>%
  filter(!is.na(.data[[enso_var]])) %>%
  arrange(date)
stopifnot(nrow(df_f14) > 0)

# Diagnostic: runs of missing days longer than 30 days (e.g. the 2007 gap)
df_f14 %>%
  mutate(gap_days = as.integer(date - lag(date)) - 1L,
         gap_start = lag(date) + 1, gap_end = date - 1) %>%
  filter(!is.na(gap_days), gap_days > 30) %>%
  select(gap_start, gap_end, gap_days) %>%
  print()

x_lims_f14 <- c(as.Date("2000-01-01"), max(df_f14$date))
y_data_max <- max(df_f14[[enso_var]], na.rm = TRUE)
y_data_min <- min(df_f14[[enso_var]], na.rm = TRUE)
stripe_y0 <- ceiling((y_data_max + 0.05) * 10) / 10   # just above the highest peak
stripe_y1 <- stripe_y0 + 0.20
y_lims_f14 <- c(floor(y_data_min * 2) / 2, stripe_y1)

# Stripe drawn from the ENSO calendar, so it is continuous across TWL gaps
stripe_f14 <- oni_daily %>%
  filter(Date >= x_lims_f14[1], Date <= x_lims_f14[2], !is.na(enso_phase))

p_f14 <- ggplot(df_f14) +
  geom_segment(
    data = stripe_f14,
    aes(x = Date, xend = Date, y = stripe_y0, yend = stripe_y1, color = enso_phase),
    linewidth = 1
  ) +
  # Line drawn from the unfiltered series so it breaks across missing stretches
  geom_line(
    data = twl_enso_photos %>% filter(mop == f14_mop, date >= as.Date("2000-01-01")),
    aes(x = date, y = .data[[enso_var]]), color = "lightblue", linewidth = 1
  ) +
  geom_point(
    data = df_f14 %>% filter(photo == 1),
    aes(x = date, y = .data[[enso_var]]),
    color = "black", size = 2.5, shape = 1, stroke = 1
  ) +
  geom_vline(
    xintercept = as.numeric(seq(x_lims_f14[1], x_lims_f14[2], by = "1 year")),
    color = "darkgray", linetype = "dashed", linewidth = 0.5
  ) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year",
               limits = x_lims_f14, expand = c(0, 0)) +
  scale_y_continuous(limits = y_lims_f14, expand = c(0, 0)) +
  scale_color_manual(
    values = c("El Niño" = "red", "La Niña" = "blue", "Neutral" = "gray"),
    breaks = c("El Niño", "La Niña", "Neutral"),
    guide = guide_legend(override.aes = list(linewidth = 3))
  ) +
  labs(x = NULL, y = paste("MOP Transect", f14_mop, enso_label), color = NULL) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    legend.text = element_text(size = 12),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid = element_blank()
  )
p_f14
save_gg(p_f14, "f14.png", width = 10, height = 6, dpi = 300)
```

```{r, eval=has_oni}
# As in the submitted manuscript: days above a fixed TWL value, all photo transects.
# -----------------------------
# f15: Extreme-event frequency by ENSO phase (selected MOPs)
# -----------------------------
selected_mops <- c(
  "D0038","D0039","D0041","D0042","D0043","D0044","D0045","D0048",
  "D0055","D0056","D0057","D0111","D0114","D0349","D0350","D0352",
  "D0353","D0363","D0373","D0374","D0377","D0379","D0386","D0397",
  "D0495","D0498","D0499","D0501","D0502","D0503","D0504","D0505",
  "D0513","D0587","D0588","D0590","D0621","D0623","D0666","D0667",
  "D0674","D0675","D0678","D0681","D0800","D0881","D0894",
  "D0896","D0906","D1022"
)

extreme_var <- "twl"   # daily (24-hour) maximum TWL
twl_thr <- params$twl_threshold_for_summary

df_f15 <- twl_enso_photos %>%
  filter(mop %in% selected_mops) %>%
  filter(!is.na(.data[[extreme_var]])) %>%
  mutate(
    enso_phase = as.character(enso_phase),
    enso_phase = ifelse(enso_phase %in% c("Neutral", "Excluded"), "Neutral", enso_phase),
    extreme_event = as.integer(.data[[extreme_var]] > twl_thr)
  )

stopifnot(nrow(df_f15) > 0)

extreme_event_by_mop_phase <- df_f15 %>%
  group_by(mop, enso_phase) %>%
  summarise(frequency_of_extremes = mean(extreme_event, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    enso_phase = factor(enso_phase, levels = c("El Niño", "La Niña", "Neutral")),
    mop = factor(mop, levels = selected_mops)
  )

p_f15 <- ggplot(extreme_event_by_mop_phase, aes(x = mop, y = frequency_of_extremes, fill = enso_phase)) +
  geom_col(position = "dodge") +
  scale_y_continuous(
    limits = c(0, 0.04),
    expand = c(0, 0),
    breaks = seq(0, 0.04, 0.01)
  ) +
  labs(
    x = "MOP Transect",
    y = paste0("Frequency of extreme TWL (daily maximum > ", twl_thr, " m)"),
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    legend.title = element_blank(),
    axis.text.x = element_text(
      angle = 90,
      hjust = 0.5,
      vjust = 0.5,
      size = 10
    ),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(margin = margin(t = 10), size = 12),
    axis.title.y = element_text(margin = margin(r = 10), size = 12),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid = element_blank(),
    legend.position = c(0.02, 0.98),
    legend.justification = c(0, 1),
    legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5)
  ) +
  scale_fill_manual(values = c("El Niño" = "red", "La Niña" = "blue", "Neutral" = "grey"))

p_f15
save_gg(p_f15, "f15.png", width = 10, height = 6, dpi = 300)
```


Additions as of 19 Sept 2026
```{r}
results_clean %>% filter(total_photos == 1) %>% count(predictor, fit_flag)
delta_aic %>% left_join(distinct(results_clean, mop, total_photos)) %>%
  count(total_photos == 1)
```

```{r}
multi_mops <- results_clean %>%
  distinct(mop, total_photos) %>%
  filter(total_photos >= 2, !mop %in% exclude_mops_default) %>%
  pull(mop)

summarise_daic <- function(d) {
  d %>% group_by(predictor) %>%
    summarise(n_transects = n(),
              median_dAIC = median(dAIC, na.rm = TRUE),
              n_better = sum(dAIC < -2, na.rm = TRUE),
              n_worse  = sum(dAIC >  2, na.rm = TRUE),
              .groups = "drop")
}

left_join(
  summarise_daic(delta_aic),
  summarise_daic(filter(delta_aic, mop %in% multi_mops)),
  by = "predictor", suffix = c("_all", "_2plus")
)
```

```{r}
one_mops <- results_clean %>% distinct(mop, total_photos) %>%
  filter(total_photos == 1) %>% pull(mop)

df_photos_max %>%
  filter(mop %in% one_mops) %>%
  pivot_longer(all_of(preds), names_to = "predictor", values_to = "x") %>%
  filter(!is.na(x), !is.na(photo)) %>%
  group_by(mop, predictor) %>%
  summarise(separated = max(x[photo == 1]) >= max(x[photo == 0]),
            .groups = "drop") %>%
  count(predictor, separated)
```

```{r}
delta_aic %>% filter(mop %in% multi_mops, predictor != baseline) %>%
  group_by(predictor) %>%
  summarise(n_pos = sum(dAIC > 0), n = n(),
            p_sign = binom.test(n_pos, n)$p.value)
```


# Bayesian prior sensitivity: Normal(0, 1.5)

```{r bayes-prior-sensitivity}
# Refit every transect model with tighter priors and compare thresholds with the
# main Normal(0, 2.5) results. Compiled once, then reused.
priors_tight <- c(
  prior(normal(0, 1.5), class = "Intercept"),
  prior(normal(0, 1.5), class = "b")
)
base_fit_tight <- brm(
  formula = photo ~ x_centered, data = example_data,
  family = bernoulli(link = "logit"), prior = priors_tight,
  chains = 2, iter = 1000, warmup = 500, seed = 123, refresh = 0
)
bayes_tight <- purrr::map_dfr(mops_keep, function(m) {
  d_m <- df_bayes %>% filter(mop == m, !is.na(photo), !is.na(x_centered))
  if (nrow(d_m) < 10 || n_distinct(d_m$photo) < 2) return(tibble())
  fit_m <- update(base_fit_tight, newdata = d_m, recompile = FALSE,
                  chains = 4, iter = 2000, warmup = 1000, seed = 123, refresh = 0,
                  control = list(adapt_delta = 0.95, max_treedepth = 12))
  qs <- quantile(threshold_draws_from_fit(fit_m, params$twl_threshold_for_summary),
                 probs = c(0.025, 0.5, 0.975), na.rm = TRUE)
  tibble(mop = m, median_tight = qs[[2]], lower_tight = qs[[1]], upper_tight = qs[[3]])
})
prior_cmp <- bayes_results %>%
  mutate(mop = as.character(mop)) %>%
  select(mop, n_flood_photos, median_main = threshold_median,
         lower_main = threshold_lower, upper_main = threshold_upper) %>%
  left_join(bayes_tight, by = "mop") %>%
  mutate(diff_median = median_tight - median_main) %>%
  mutate(across(where(is.double), ~ round(.x, 2)))
old_opts <- options(width = 220)
print(prior_cmp, n = Inf, width = Inf)
prior_cmp %>%
  mutate(group = if_else(n_flood_photos >= 8, "8+ photos", "2-7 photos")) %>%
  group_by(group) %>%
  summarise(n = n(), median_diff = median(diff_median),
            max_abs_diff = max(abs(diff_median)), .groups = "drop") %>%
  print(width = Inf)
options(old_opts)
readr::write_csv(prior_cmp, here::here(params$tables_dir, "bayes_prior_sensitivity.csv"))
```

# Numbers for the reviewer responses

```{r response-numbers}
# Printed at full width so nothing is hidden in the HTML log.
old_opts <- options(width = 220)

# 1. Headline-predictor thresholds with delta-method intervals
results_main %>%
  filter(predictor == hp) %>%
  select(mop, total_photos, n_events, threshold, lower_95, upper_95, slope, p_value) %>%
  print(n = Inf, width = Inf)

# 2. Balanced documentation thresholds: the predictor value at which the fitted
#    probability equals the transect's overall photo frequency (crossing point of
#    the photo and no-photo distributions; equal class weights). Delta-method
#    interval, taking the photo frequency as fixed. Reported alongside the
#    documentation threshold (fitted probability = 0.5).
balanced_tbl <- purrr::map_dfr(mops_main, function(m) {
  d   <- df_photos_max %>% filter(mop == m, !is.na(.data[[hp]]), !is.na(photo))
  mod <- glm(as.formula(paste("photo ~", hp)), data = d, family = binomial)
  b   <- unname(coef(mod)); V <- vcov(mod)
  k   <- qlogis(mean(d$photo))
  balanced_thr <- (k - b[1]) / b[2]
  g   <- c(-1 / b[2], -(k - b[1]) / b[2]^2)
  se  <- sqrt(as.numeric(t(g) %*% V %*% g))
  doc_thr <- -b[1] / b[2]
  x1  <- d[[hp]][d$photo == 1]
  tibble(mop = m, n_photos = sum(d$photo), n_days = nrow(d),
         photo_rate = mean(d$photo),
         documentation_threshold = doc_thr,
         balanced_threshold = balanced_thr,
         balanced_lower = balanced_thr - 1.96 * se,
         balanced_upper = balanced_thr + 1.96 * se,
         photos_above_balanced = sum(x1 > balanced_thr),
         photos_above_documentation = sum(x1 > doc_thr))
}) %>%
  mutate(across(c(documentation_threshold, balanced_threshold,
                  balanced_lower, balanced_upper), ~ round(.x, 2)),
         photo_rate = signif(photo_rate, 3))
print(balanced_tbl, n = Inf, width = Inf)
balanced_tbl %>%
  summarise(median_balanced = median(balanced_threshold),
            min_balanced = min(balanced_threshold),
            max_balanced = max(balanced_threshold),
            median_documentation = median(documentation_threshold),
            photos = sum(n_photos), above_balanced = sum(photos_above_balanced),
            above_documentation = sum(photos_above_documentation)) %>%
  print(width = Inf)
readr::write_csv(balanced_tbl, here::here(params$tables_dir, "balanced_thresholds_by_mop.csv"))

# 3. Delta versus Fieller intervals, all columns
if (exists("cmp")) print(cmp %>% mutate(across(where(is.numeric), ~ round(.x, 3))), n = Inf, width = Inf)

# 4. AR GLMM odds ratios behind Fig. 12
if (exists("ar_glmm_results")) print(ar_glmm_results, n = Inf, width = Inf)

# 5. Photo transect-days that could not be matched to the panel
unmatched_path <- here::here("output", "photos_unmatched.csv")
if (file.exists(unmatched_path)) print(readr::read_csv(unmatched_path, show_col_types = FALSE), n = Inf, width = Inf)

options(old_opts)
```


```{r session-info, echo=FALSE}
sessionInfo()
```

